simulation_study_spline

### 2025-02-16 R/simulation_study/examples/illustrative_trajectories.Rmd
update .gitignore

```{r}
fomega_cubic=function(cov_data,a=1,D0=0.3,D1=0.8,sigma_theta=pi/8,sigma_D=0.2,b=5){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

@@

@@

![Simulated Trajectories](sim_fjords_cubic_attractive_sub.gif)
```{r}

# Add covariates to simulation -----------

signed_angle <- function(u, v) {
  #Compute signed angle in [-pi,pi] that rotates first vector into second vector 
  # as in 
  # https://math.stackexchange.com/questions/529555/signed-angle-between-2-vectors
  u <- matrix(u, ncol = 2)
  v <- matrix(v, ncol = 2)
  if (nrow(u) != nrow(v)) stop("u and v must have the same number of 
                                  rows")
  result <- as.numeric(atan2(v[,2], v[,1]) - atan2(u[,2], u[,1]))
  ind1 <- which(result > pi)
  ind2 <- which(result <= -pi)
  result[ind1] <- result[ind1] - 2*pi
  result[ind2] <- result[ind2] + 2*pi
  return(result) 
}

add_covs=function(data) {
  #data has at least columns "time","y1" "y2", "p1","p2"
  #this function new data frame with three more columns "theta", "DistanceShore" and "ExpShore"

  new_data=data
  #observed deviation angles
  n=length(new_data$time)

  new_data$theta=rep(NA,n)

  for (id in unique(new_data$ID)) {

    #filter data with specific id
    sub_ind=(new_data$ID==id)
    sub_data=new_data[sub_ind,]
    n_sub=length(sub_data$time)

    #time steps
    dtimes=sub_data[2:n_sub,"time"]-sub_data[1:(n_sub-1),"time"]

    #step lengths
    dx=sub_data[2:n_sub,"y1"]-sub_data[1:(n_sub-1),"y1"]
    dy=sub_data[2:n_sub,"y2"]-sub_data[1:(n_sub-1),"y2"]

    #matrix of empirical velocity
    vexp_df=cbind(dx/dtimes,dy/dtimes)

    #matrix of normal vectors
    normal=as.matrix(sub_data[2:n_sub,c("y1","y2")]-sub_data[2:n_sub,c("p1","p2")])

    #angle between velocity and normal vector
    theta_coast=signed_angle(normal,vexp_df)

    #adjust lengths 
    theta_coast=c(theta_coast,1)

    new_data[sub_ind,"theta"]=theta_coast
  }

  #distance to shore
  Dshore=sqrt((new_data$y1-new_data$p1)^2+(new_data$y2-new_data$p2)^2)
  new_data$DistanceShore=Dshore
  new_data$ExpShore=1/Dshore

  return(new_data)
}

```{r}
data_sim_fjords_cubic_attractive_sub=add_covs(data_sim_fjords_cubic_attractive_sub)
data_sim_fjords_cubic_attractive_sub$angular_velocity=
  c(NA,diff(data_sim_fjords_cubic_attractive_sub$theta)/
      diff(data_sim_fjords_cubic_attractive_sub$time))

ggplot(data=data_sim_fjords_cubic_attractive_sub)+geom_point(aes(x=time,y=DistanceShore,col=ID))+
  geom_line(aes(x=time,y=DistanceShore,col=ID),size=1,alpha=0.2)

ggplot(data=data_sim_fjords_cubic_attractive_sub)+geom_point(aes(x=time,y=angular_velocity,col=ID))+
  geom_line(aes(x=time,y=angular_velocity,col=ID),size=1,alpha=0.2)
```

@@

fomega_cubic_fjords_repulsive=function(cov_data) {
  fomega_cubic(cov_data,a=40,D0=0.3,D1=0.8,sigma_theta=pi/10,sigma_D=0.3,b=0)
}

@@

![Simulated Trajectories](sim_fjords_cubic_repulsive_sub.gif)
```{r}
data_sim_fjords_cubic_advanced_gaussian_sub=add_covs(data_sim_fjords_cubic_advanced_gaussian_sub)
data_sim_fjords_cubic_advanced_gaussian_sub$angular_velocity=
  c(NA,diff(data_sim_fjords_cubic_advanced_gaussian_sub$theta)/
      diff(data_sim_fjords_cubic_advanced_gaussian_sub$time))

ggplot(data=data_sim_fjords_cubic_advanced_gaussian_sub)+geom_point(aes(x=time,y=DistanceShore,col=ID))+
  geom_line(aes(x=time,y=DistanceShore,col=ID),size=1,alpha=0.2)

ggplot(data=data_sim_fjords_cubic_advanced_gaussian_sub)+geom_point(aes(x=time,y=angular_velocity,col=ID))+
  geom_line(aes(x=time,y=angular_velocity,col=ID),size=1,alpha=0.2)
```

@@

```{r simulate hanks reflected in rectangle,cache=TRUE}
```

@@

  ggplot() +
    geom_sf(data = geometries %>% filter(domain == !!domain), aes(geometry = geometry), fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
    theme_minimal() +
    xlab("x") + ylab("y") +
    theme(legend.position = "none") +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 16),
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )
}

>>>>

```{r}
fomega_cubic=function(cov_data,a=1,b=5,D0=0.3,D1=0.8,sigma_theta=pi/8,sigma_D=0.3){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

@@

Finally, we compare this to Hanks simulation technique for reflected SDE.

@@

![Simulated Trajectories](sim_fjords_cubic_attractive_sub.gif)

@@

fomega_cubic_fjords_repulsive=function(cov_data) {
  fomega_cubic(cov_data,a=1,D0=0.3,D1=0.8,sigma_theta=pi/8,sigma_D=0.3,b=0)
}

@@

![Simulated Trajectories](sim_fjords_cubic_repulsive_sub.gif)

@@

@@

  ggplot() +
    geom_sf(data = geometries %>% filter(domain == !!domain), aes(geometry = geometry), fill = "grey", inherit.aes = FALSE) +
    coord_sf(datum = st_crs("+init=EPSG:32626 +units=km")) +
    geom_path(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = data_filtered, aes(x = y1, y = y2, col = ID), size = 0.1) +
    geom_point(data = x0, aes(x = x1, y = x2), shape = 3, size = 2, col = "red") +
     scale_color_viridis_d(labels=c("A1","A2","A3","A4","A5","A6"))+
    theme_minimal() + 
    xlab("x") + ylab("y") +
    theme(legend.position = "none") +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 18),
    axis.title.y = element_text(angle = 0)  # Rotate the y-axis label and adjust horizontal justification
  )
}

### 2025-02-16 R/simulation_study/local/splines.Rmd
update .gitignore

```{r}
library(here)
library(sf)
library(smoothSDE)
library(doParallel)
library(ggplot2)
library(dplyr)
library(mgcv)
library(plotly)

#get functions to simulate trajectories
source(file.path(here("R","simulation_study_CRCVM","CVM_functions.R")))

# Simple framework : no random effects

## Define hyperparameters

```{r hyperparameters}
N_ID=6

TMAX=12
DELTA=1/60/6

SP_DF=c(5,5)

TAU_0=1.3
NU_0=4.6

DMIN=0.5
DMAX=1

PAR0_RACVM=c(0,0,1,1,0)

SIGMA_OBS_LOW=0.01
SIGMA_OBS_HIGH=0.04

BY_LF=5
BY_HF=1

D_LOW=0.05
D_UP=3

A=1
D0=0.3
D1=1.2
SIGMA_THETA=pi/6
SIGMA_D=1
B=1

```{r initial positions}

seed=42
set.seed(seed)
# Set the path to the directory containing the data
greenland_data_path <- here("Data","preprocessed_data","greenland")

#get the land geometry from shapefile
border<-st_read(file.path(greenland_data_path,"updated_scoresby_sound_utm.shp"))
border <- st_transform(border, crs = "+init=EPSG:32626 +units=km")

shrunk_border<-st_read(file.path(greenland_data_path,"shrunk_updated_scoresby_sound_utm.shp"))
shrunk_border <- st_transform(border, crs = "+init=EPSG:32626 +units=km")
v0=c(0,0)

#generate random initial points
x0=matrix(rep(NA,N_ID*2),ncol=2)
colnames(x0)=c("x1","x2")

i=1
while (i<=N_ID) {
  #choose location uniformly in the map
  x=c(runif(1,min=430,max=500),runif(1,min=7760,max=7900))
  if (!(is_in_land(st_point(x),border))) {
    p=nearest_boundary_point(st_point(x),border)
    Dshore=(x[1]-p[1])^2+(x[2]-p[2])^2
    #keep it as initial position if it is at least 50 metres away from the shore
    if (Dshore>DMIN && Dshore<DMAX) {
      x0[i,]=x
      i=i+1
    }
  }
}

## Define SDE parameters

```{r model parameters}
# Time steps ----------------

#high frequency time points
times=seq(0,TMAX,by=DELTA)

n_obs=length(times)-1

# Defintion of smooth parameter omega ----------

fomega_cubic=function(cov_data,a,b,D0,D1,sigma_D,sigma_theta){
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  if (is.null(Dshore)){
    Dshore=1/cov_data$Ishore
  }

  a*theta*(theta-pi/2)*(theta+pi/2)*exp(-Dshore/D0)/Dshore+
      b*(exp(-1/2*(((theta+pi/2/sqrt(3))/sigma_theta)^2+((Dshore-D1)/sigma_D)^2))-
           exp(-1/2*(((theta-pi/2/sqrt(3))/sigma_theta)^2+((Dshore-D1)/sigma_D)^2)))
}

fomega=function(cov_data) {
  fomega_cubic(cov_data,A,B,D0,D1,SIGMA_D,SIGMA_THETA)
}

ftau_constant=function(cov_data) {
  TAU_0
}

fnu_constant=function(cov_data) {
  NU_0
}

## Simulate true trajectories

```{r,cache=TRUE}

# Generate samples ---------------

cores=detectCores()
cl <- makeCluster(6) #not to overload your computer
registerDoParallel(cl)

data=foreach (i=1:N_ID,.combine='rbind',.packages=c("progress","MASS","sf","mgcv")) %dopar% {

  set.seed((seed-1)*N_ID+i)

  res=sim_CRCVM(ftau=ftau_constant,fomega=fomega,fnu=fnu_constant,
                      log_sigma_obs=NULL,v0=v0,x0=x0[i,],times=times,
                land=shrunk_border,verbose=FALSE,save_omega = TRUE)

  data_sim=res$sim
  data_sim$ID=factor(rep(i,length(data_sim$y1)))
  data_shore=res$shore

  data_sim=cbind(data_sim,data_shore)
  data_sim
}

#stop cluster
stopCluster(cl)

## Remove trajectories that reached land

# Points that reached land ---------------
count=0
remove_ID=c()
for (id in unique(data$ID)) {
  id_data=data[data$ID==id,]
  if (nrow(id_data) < n_obs) {
    count=count+1
    remove_ID=c(remove_ID,id)
    cat("ID",id,"reached land","\n",sep=" ")
  }
}

cat(count/N_ID*100,"percent of the trajectories reached land","\n")

data_sub=data %>%  filter(!(ID %in% remove_ID))

## Add noise to get observed trajectories

```{r}
# add noise
low_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_LOW^2,2)))
high_noise=rmvn(n_obs,rep(0,2),diag(rep(SIGMA_OBS_HIGH^2,2)))
observed_data_le=data_sub[,c("y1","y2","time","ID")]
observed_data_le[,c("y1","y2")]=data_sub[,c("y1","y2")]+low_noise
observed_data_he=data_sub[,c("y1","y2","time","ID")]
observed_data_he[,c("y1","y2")]=data_sub[,c("y1","y2")]+high_noise

#subsample
data_lf_he=observed_data_he[seq(1,length(data_sub$time),by=BY_LF),]
data_lf_le=observed_data_le[seq(1,length(data_sub$time),by=BY_LF),]
data_hf_le=observed_data_le[seq(1,length(data_sub$time),by=BY_HF),]
data_hf_he=observed_data_he[seq(1,length(data_sub$time),by=BY_HF),]

# Save plot of the trajectories ------------

plot=ggplot()+geom_sf(data=border$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_point(data=data_sub,mapping=aes(y1,y2,col=ID),size=0.1,alpha=0.2,shape=2)+
  geom_point(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_path(data=data_lf_he,mapping=aes(y1,y2,col=ID),size=0.1)+
  geom_point(data = data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

ggsave(filename="plot_trajectories.pdf",plot=plot,width=10,height=5)

## Compute covariates from observations

# Add covariates to data -----------

signed_angle <- function(u, v) {
  #Compute signed angle in [-pi,pi] that rotates first vector into second vector 
  # as in 
  # https://math.stackexchange.com/questions/529555/signed-angle-between-2-vectors
  u <- matrix(u, ncol = 2)
  v <- matrix(v, ncol = 2)
  if (nrow(u) != nrow(v)) stop("u and v must have the same number of 
                                  rows")
  result <- as.numeric(atan2(v[,2], v[,1]) - atan2(u[,2], u[,1]))
  ind1 <- which(result > pi)
  ind2 <- which(result <= -pi)
  result[ind1] <- result[ind1] - 2*pi
  result[ind2] <- result[ind2] + 2*pi
  return(result) 
}

  # Function to calculate rolling average
rolling_average <- function(data, window) {
  n <- nrow(data)
  result <- matrix(NA, n, ncol(data))  # Initialize result matrix with NA
  colnames(result) <- colnames(data)

  for (i in seq_len(n)) {
    # Define the range for the current window
    start <- max(1, i - floor(window / 2))
    end <- min(n, i + floor(window / 2))
    result[i, ] <- colMeans(data[start:end, , drop = FALSE])  # Compute column means
  }
  return(result)
}

add_covs_parallel <- function(data,add="all",window=1,n_cores = parallel::detectCores() - 1) {
  # Setup parallel backend
  cl <- makeCluster(n_cores)
  registerDoParallel(cl)

  # Split data by ID
  ids <- unique(data$ID)

  # Process each ID in parallel
  results <- foreach(id = ids, .combine = rbind, .packages = c("sf"),
                     .export = c("nearest_boundary_points", "is_in_land", "signed_angle","border",
                                 "D_LOW","D_UP","rolling_average")  ) %dopar% {
    # Filter data for the current ID
    sub_data <- data[data$ID == id, ]
    n_sub <- nrow(sub_data)

    # Time steps
    dtimes <- sub_data[2:n_sub, "time"] - sub_data[1:(n_sub - 1), "time"]

    # Step lengths
    dx <- sub_data[2:n_sub, "y1"] - sub_data[1:(n_sub - 1), "y1"]
    dy <- sub_data[2:n_sub, "y2"] - sub_data[1:(n_sub - 1), "y2"]

    # Empirical velocity
    vexp_df <- cbind(dx / dtimes, dy / dtimes)

    vexp_avg <- rolling_average(vexp_df, window)

    # Nearest points on shore
    sub_data <- cbind(sub_data, nearest_boundary_points(as.matrix(sub_data[, c("y1", "y2")]), border))

    # Normal vectors
    normal <- as.matrix(sub_data[2:n_sub, c("y1", "y2")] - sub_data[2:n_sub, c("p1", "p2")])

    # Angle between velocity and normal vector
    theta_coast <- signed_angle(normal, vexp_avg)
    theta_coast <- c(theta_coast, 1)  # Adjust length

    # Initialize DistanceShore
    Dshore <- rep(0, n_sub)

    for (i in 1:n_sub) {
      y1 <- sub_data[i, "y1"]
      y2 <- sub_data[i, "y2"]
      if (!is_in_land(st_point(c(y1, y2)), border)) {
        p1 <- sub_data[i, "p1"]
        p2 <- sub_data[i, "p2"]
        Dshore[i] <- sqrt((y1 - p1)^2 + (y2 - p2)^2)
      }
    }

    Dshore_avg=as.vector(rolling_average(matrix(Dshore,ncol=1),window))

    # Calculate Ishore
    Ishore_avg <- ifelse(Dshore_avg < D_LOW, 1 / D_LOW, ifelse(Dshore_avg > D_UP, 0, 1 / Dshore_avg))

    # Add columns to sub_data
    if (add=="all") {
      sub_data$theta <- theta_coast
      sub_data$DistanceShore <- Dshore_avg
      sub_data$Ishore <- Ishore_avg
    }
    else if (add=="theta") {
      sub_data$theta <- theta_coast
    }
    else if (add=="DistanceShore") {
      sub_data$DistanceShore <- Dshore
      sub_data$Ishore <- Ishore
        }
  else {
    stop("Options for argument add are \"all\",\"DistanceShore\" and \"theta\"")
  }

    return(sub_data)
  }

  # Stop the cluster
  stopCluster(cl)

  return(results)
}

```{r}
#colnames(data)[8]="DistanceShore"
data_lf_he=add_covs_parallel(data_lf_he,window=3)
data_hf_le=add_covs_parallel(data_hf_le,window=3)
data_lf_le=add_covs_parallel(data_lf_le,window=3)
data_hf_he=add_covs_parallel(data_hf_he,window=3)
```

plot_Dshore=ggplot(data=data_lf_he)+geom_point(aes(x=time,y=DistanceShore,col=ID),size=1,alpha=0.5)+
  geom_line(aes(x=time,y=DistanceShore,col=ID),size=1,alpha=0.2)+facet_wrap(~ID)+
  theme_minimal()

plot_Dshore

ggplot(data_lf_he)+geom_histogram(aes(x=DistanceShore),bins=30)+facet_wrap(~ID)+theme_minimal()
quantile(data_lf_he$DistanceShore)
```

data_lf_he$angular_velocity=as.vector(sapply(unique(data_sub$ID),function(id){
  c(NA,diff(data_lf_he[data_lf_he$ID==id,"theta"])/diff(data_lf_he[data_lf_he$ID==id,"time"]))}))

plot_angular_velocity=ggplot(data=data_lf_he)+geom_point(aes(x=time,y=angular_velocity,col=ID),size=1,alpha=0.5)+
  geom_line(aes(x=time,y=angular_velocity,col=ID),size=1,alpha=0.2)+facet_wrap(~ID)+
  theme_minimal()
plot_angular_velocity

ggplot(data_lf_he)+geom_histogram(aes(x=angular_velocity),bins=30)+facet_wrap(~ID)+theme_minimal()
quantile(data_lf_he$angular_velocity,na.rm=TRUE)

```{r}
data_lf_he$speed=as.vector(sapply(unique(data_sub$ID),function(id) {c(NA,sqrt(diff(data_lf_he[data_lf_he$ID==id,"y1"])^2+diff(data_lf_he[data_lf_he$ID==id,"y2"])^2)
                                                                      /diff(data_lf_he[data_lf_he$ID==id,"time"]))}))

ggplot(data_lf_he)+geom_histogram(aes(x=speed),bins=30)+facet_wrap(~ID)+theme_minimal()
quantile(data_lf_he$speed,na.rm=TRUE)
```
```{r}
ggplot(data_lf_he)+geom_histogram(aes(x=theta),bins=30)+facet_wrap(~ID)+theme_minimal()
```

## Estimate CTCRW

### low frequency high error

@@

#fit
ctcrw_lf_he$fit(method="BFGS")
```
```{r}
ctcrw_lf_he$get_all_plots()
exp(as.list(ctcrw_lf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,nu=~1,omega=~1)

@@

#fit
ctcrw_lf_le$fit(method="BFGS")
```
```{r}
ctcrw_lf_le$get_all_plots()
exp(as.list(ctcrw_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

ctcrw_hf_le<- SDE$new(formulas = formulas,data =data_hf_le,type = "CTCRW",
                    response = c("y1","y2"),par0 = c(0,0,1,1),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.03)))

#fit
ctcrw_hf_le$fit(method="BFGS")
```
```{r}
ctcrw_hf_le$get_all_plots()

@@

@@

#fit
ctcrw_hf_he$fit(method="BFGS")
```
```{r}
ctcrw_hf_he$get_all_plots()

@@

```{r}
# Splines smooth for omega -------------------- 
n <- 100000
SP_DF=5

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.1,2)
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

@@

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=SP_DF),knots=list(theta=seq(-pi,pi,len=SP_DF),
                                                       DistanceShore=seq(0.02,7,len=SP_DF)))

ExpShore=1/DistanceShore

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=SP_DF),knots=list(theta=seq(-pi,pi,len=SP_DF),
                                                  ExpShore=seq(1/8,1/0.2,len=SP_DF)))

fomega_cubic_splines=function(cov_data) {

  if (is.null(cov_data$ExpShore)){
    omega=predict(m1,newdata=cov_data)
  }
  else if (is.null(cov_data$DistanceShore)){
    omega=predict(m2,newdata=cov_data)
  }

  return(as.numeric(omega))
}

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=SP_DF,bs=c("cs")))

crcvm_splines_lf_he<- SDE$new(formulas = formulas,data = data_lf_he,type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

crcvm_splines_lf_he$update_map(list("log_lambda"=factor(c(rep(NA,2)))))
crcvm_splines_lf_he$update_lambda(m1$sp)
#fit
crcvm_splines_lf_he$fit(method="BFGS")
```

@@

crcvm_splines_lf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

estimate_parametric_omega=function(model) {

@@

    Dshore=seq(from=quantile(data$DistanceShore,0.5),to=quantile(data$DistanceShore,0.95),length.out=30)
    theta=seq(from=-pi,to=pi,length.out=30)
    grid<- as.data.frame(expand.grid(Dshore,theta))
    colnames(grid) <- c("DistanceShore","theta")
    grid$ID=1

@@

    fit_optim <- optim(par = c(0.1,2),
                       fn = function(params) {

                        # Compute predicted z values
                        z_pred <- fomega_cubic(data,A,B,params[1],params[2],SIGMA_D,SIGMA_THETA)

@@

    est_par=fit_optim$par

    # Compute the predicted values using the starting parameters
    z_pred_starting <- fomega_cubic(grid,A,B,est_par[1],est_par[1],SIGMA_D,
                                    SIGMA_THETA)

    # Reshape z_pred_starting to a matrix for plotting
    z_pred_matrix <- matrix(z_pred_starting, nrow = length(Dshore), ncol = length(theta))

@@

estimate_parametric_omega(crcvm_splines_lf_he)

@@

crcvm_splines_lf_le<- SDE$new(formulas = formulas,data = data_lf_le[data_lf_le$ID %in% c(1:6),],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

@@

crcvm_splines_lf_le$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

exp(as.list(crcvm_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

crcvm_splines_hf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

estimate_parametric_omega(crcvm_splines_hf_he)

@@

## Estimation with fixed parametric omega function

We will fix the new parameters that are introduced in the model, and estimate the others, that is $\tau$ and $\nu$.
These new parameters might be chosen based on the observed distance to shore and observed angular velocity in the data.

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_lf_he$time)),dim=c(2,2,length(data_lf_he$time)))
# Estimate from simulated data ------------------

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)

crcvm_cubic_lf_he<- SDE$new(formulas = formulas,
                            data = data_lf_he,type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

#fit
crcvm_cubic_lf_he$fit(method="BFGS")
```

crcvm_cubic_lf_he$get_all_plots(show_CI="simultaneous")
exp(as.list(crcvm_cubic_lf_he$tmb_rep(),what="Estimate")$log_sigma_obs)

We are unable to estimate $\tau$ while fixing $\omega$. $\tau$ is totally biased and underestimated, even more than when we neglected the constraints and used a simple CTCRW..

```{r}
H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_lf_le$time)),dim=c(2,2,length(data_lf_le$time)))
# Estimate from simulated data ------------------

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)

crcvm_cubic_lf_le<- SDE$new(formulas = formulas,data = data_lf_le[data_lf_le$ID %in% c(1:4),],
                              type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_lf_le$fit(method="BFGS")
```
```{r}

crcvm_cubic_lf_le$get_all_plots(show_CI="simultaneous")
exp(as.list(crcvm_cubic_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

```
Let's now try with high frequency data.

```{r}
H=array(rep(SIGMA_OBS_HIGH^2*diag(2),length(data_hf_he$time)),dim=c(2,2,length(data_hf_he$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)

crcvm_cubic_hf_he<- SDE$new(formulas = formulas,data = data_hf_he[data_hf_he$DistanceShore>0.01 & data_hf_he$ID %in% c(1:6),],type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 =c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

crcvm_cubic_hf_he$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_he$get_all_plots()
exp(as.list(crcvm_cubic_hf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```

H=array(rep(SIGMA_OBS_LOW^2*diag(2),length(data_hf_le$time)),dim=c(2,2,length(data_hf_le$time)))

formulas <- list(tau=~1,
                 nu=~1,a=~1,b=~1,D0=~1,D1=~1,sigma_D=~1,sigma_theta=~1)

crcvm_cubic_hf_le<- SDE$new(formulas = formulas,data = data_hf_le[data_hf_le$ID %in% c(1:4),],
                            type = "CRCVM_SSM",
                    response = c("y1","y2"),par0 = c(1,1,A,B,D0,D1,SIGMA_D,SIGMA_THETA),
                    fixpar=c("a","b","D0","D1","sigma_D","sigma_theta"),
                    other_data=list("H"=H))

#fit
crcvm_cubic_hf_le$fit(method="BFGS")
```

```{r}
crcvm_cubic_hf_le$get_all_plots()
exp(as.list(crcvm_cubic_hf_le$tmb_rep(),what="Estimate")$log_sigma_obs)
```

```{r}
ggplot()+geom_point(data=data_hf_le,aes(time,theta),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,theta),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)

```
```{r}
data_hf_le$omega=fomega(data_hf_le[,c("theta","DistanceShore")])

ggplot()+geom_point(data=data_hf_le,aes(time,omega),col="blue",alpha=0.5,size=0.1)+
  geom_point(data=data,aes(time,omega),col="red",alpha=0.2,size=0.1)+facet_wrap(~ID)
```

>>>>

# Set up for simulation study
```{r}
source(file.path(here("R","simulation_study_CRCVM","estimation_local","set_up.R")))  
```

# Estimate CTCRW

## low frequency high error

@@

#fit
ctcrw_lf_he$fit(method="BFGS")
```

```{r}
ctcrw_lf_he$get_all_plots()
exp(as.list(ctcrw_lf_he$tmb_rep(),what="Estimate")$log_sigma_obs)
```

@@

```{r fit sde model with splines,cache=TRUE}
formulas <- list(mu1=~1,mu2=~1,tau=~1,nu=~1,omega=~1)

@@

#fit
ctcrw_lf_le$fit(method="BFGS")
```

```{r}
ctcrw_lf_le$get_all_plots()
exp(as.list(ctcrw_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

ctcrw_hf_le<- SDE$new(formulas = formulas,data =data_hf_le,type = "CTCRW",
                    response = c("y1","y2"),par0 = c(0,0,1,1),fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.03)))
#fit
ctcrw_hf_le$fit(method="BFGS")
```

```{r}
ctcrw_hf_le$get_all_plots()

@@

```{r fit sde model with splines,cache=TRUE}
formulas <- list(mu1=~1,mu2=~1,tau=~1,nu=~1)

@@

#fit
ctcrw_hf_he$fit(method="BFGS")
```

```{r}
ctcrw_hf_he$get_all_plots()

@@

```{r}
# Splines smooth for omega -------------------- 
n <- 100000
SP_DF=c(5,5)

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.1,2)
samples=data.frame(BoundaryAngle=theta,BoundaryDistance=DistanceShore)

@@

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=SP_DF,bs=c("cs","cs")))

#visualize
vis.gam(m1);title("tensor product")

@@

formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(observed_BoundaryAngle,observed_BoundaryDistance,k=SP_DF,bs=c("cs")))

crcvm_splines_lf_he<- SDE$new(formulas = formulas,data = data_list[['data_lf_he']],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.05)))

crcvm_splines_lf_he$update_map(list("log_lambda"=factor(c(rep(NA,2)))))
crcvm_splines_lf_he$update_lambda(m1$sp)

#fit
crcvm_splines_lf_he$fit(method="BFGS")
```

@@

crcvm_splines_lf_he$get_all_plots(show_CI="pointwise")

@@

estimate_parametric_omega=function(model,boundary_distance_name,boundary_angle_name,
                                   fixed_a=NULL,probs=c(0.5,0.95)) {

@@

    Dshore=seq(from=quantile(data$observed_BoundaryDistance,probs[1]),
               to=quantile(data$observed_BoundaryDistance,probs[2]),
               length.out=30)
    theta=seq(from=-pi,to=pi,length.out=30)
    grid<- as.data.frame(expand.grid(Dshore,theta))
    colnames(grid) <- c(boundary_distance_name,boundary_angle_name)
    grid$ID=1

@@

    if (is.null(fixed_a)) {

      fit_optim <- optim(par = log(c(1,0.5,0.1,2,0.5,pi/6)),
                       fn = function(params) {

                        # Compute predicted z values
                        z_pred <- fomega_cubic(data,
                                         boundary_distance_name,
                                         boundary_angle_name,
                                         exp(params[1]),
                                         exp(params[2]),exp(params[3]),
                                         exp(params[4]),exp(params[5]),
                                         exp(params[6]))

@@

    } 
    else {
    fit_optim <- optim(par = log(c(0.5,0.1,2,0.5,pi/6)),
                       fn = function(params) {

                        # Compute predicted z values
                        z_pred <- fomega_cubic(data,
                                         boundary_distance_name,
                                         boundary_angle_name,
                                         fixed_a,exp(params[1]),
                                         exp(params[2]),exp(params[3]),
                                         exp(params[4]),exp(params[5]))

    }
    est_par=exp(fit_optim$par)

    # Compute the predicted values using the starting parameters
    if (is.null(fixed_a)) {

      z_pred <- fomega_cubic(grid,
                       boundary_distance_name,
                      boundary_angle_name,
                      est_par[1],
                      est_par[2],est_par[3],
                      est_par[4],est_par[5],
                      est_par[6])

    } else {
    z_pred <- fomega_cubic(grid,boundary_distance_name,
                      boundary_angle_name,
                     fixed_a,est_par[1],est_par[2],
                      est_par[3],est_par[4],est_par[5])
    }

    # Reshape z_pred_starting to a matrix for plotting
    z_pred_matrix <- matrix(z_pred, nrow = length(Dshore), ncol = length(theta))

@@

estimate_parametric_omega(crcvm_splines_lf_he,probs=c(0.75,0.95),
                          boundary_distance_name = "observed_BoundaryDistance",
                          boundary_angle_name = "observed_BoundaryAngle")

@@

crcvm_splines_lf_le<- SDE$new(formulas = formulas,data = data_list[["data_lf_he"]],type = "RACVM_SSM",
                    response = c("y1","y2"),par0 = PAR0_RACVM,fixpar=c("mu1","mu2"),
                    other_data=list("log_sigma_obs0"=log(0.01)))

@@

crcvm_splines_lf_le$get_all_plots(show_CI="pointwise")

exp(as.list(crcvm_splines_lf_le$tmb_rep(),what="Estimate")$log_sigma_obs)

@@

```{r}
crcvm_splines_hf_he$get_all_plots(show_CI="pointwise",true_smooths=list("omega"=fomega))

@@

```{r}
estimate_parametric_omega(crcvm_splines_hf_he)
```

@@
